---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pgIRT
`pgIRT` is an **R** package that implements Item Response Theory (IRT) model with Polya-Gamma data augmentation and EM Algorithm. The function is available for the data containing K >= 2 response category (binary ~ K-multinomial) and having different categories across items. In addition, the implementation includes the dynamic IRT model. The algorithm here is based on Goplerud (2019). This package also utilizes the parametric bootstrap method proposed by Lewis and Poole (2004) to estimate confidence interval.
<!-- badges: start -->
<!-- badges: end -->



## Installation

You can install the development version of pgIRT from [GitHub](https://github.com) with:

``` r
remotes::install_github("vkyo23/pgIRT")
```
## Usage

### Binary IRT

Monte Carlo simulation:

```{r}
library(pgIRT)
require(dplyr)
require(tidyr)
set.seed(1)

# Number of individuals and items
I <- 100 
J <- 1000

# DGP
theta_true <- seq(-2, 2, length = I)
alpha_true <- rnorm(J)
beta_true <- rnorm(J)
Y_star <- cbind(1, theta_true) %*% rbind(alpha_true, beta_true)
Y <- matrix(rbinom(I * J, 1, plogis(Y_star)), I, J)
```

Function `pgIRT` implements IRT. Please note that the elements of an input matrix should start from 1, not 0. Then, I convert 0 into 2.

```{r}
Y[Y == 0] <- 2 
pgfit <- pgIRT(Y,
               model = "default",
               constraint = which.max(theta_true))

cor(pgfit$parameter$theta, theta_true)
```

### Dynamic binary IRT

Using `Rehnquist` data from `MCMCpack`. To introduce some auxiliary function of `pgIRT`, I first construct long-format dataframe.

```{r}
# Convert into long-format dataframe
data(Rehnquist, package = "MCMCpack")
long_df <- Rehnquist %>% 
  mutate(rcid = row_number()) %>% 
  select(-term) %>% 
  mutate(time = time - min(time)) %>% 
  pivot_longer(-c(time, rcid))
jname <- tibble(name = unique(long_df$name)) %>% 
  mutate(judge_id = row_number())
long_df <- long_df %>% 
  left_join(jname, by = "name")
head(long_df) %>% 
  knitr::kable()
```


```{r}
# Generating roll-call matrix from long-format dataframe
mat_dyn <- make_rollcall(long_df, 
                         unit_id = "judge_id", 
                         bill_id = "rcid",
                         vote_col = "value")
rownames(mat_dyn) <- jname$name
mat_dyn[mat_dyn == 0] <- 2
constraint_dyn <- jname$judge_id[jname$name == "Thomas"]

# Options for dynamic estimation
dyn_ops <- make_dyn_options(long_df,
                            unit_id = "judge_id",
                            bill_id = "rcid",
                            time_id = "time",
                            vote_col = "value")
fit_dyn <- pgIRT(mat_dyn,
                 model = "dynamic",
                 constraint = constraint_dyn,
                 dyn_options = dyn_ops,
                 verbose = 10)
```

To compute confidence interval, run bootstrap via `pgIRT_boot`:

```{r}
# Bootstrap
boot <- pgIRT_boot(fit_dyn, boot = 100, verbose = 20)

summary(boot, parameter = "theta", ci = .95)
```


### Multinomial IRT

Using a simulated multinomial response data (`m_data_dyn`) data.

```{r}
data(m_data_dyn)

m_mlt_d <- make_rollcall(m_data_dyn,
                         unit_id = "unit",
                         bill_id = "bill",
                         vote_col = "vote") %>% 
  clean_rollcall()

fit_mlt <- pgIRT(m_mlt_d,
                 model = "default",
                 constraint = 1,
                 verbose = 20)

summary(fit_mlt)$theta
```


### Dynamic multinomial IRT

Using the simulated response data and a simulated matching bill data `sim_match` for dynamic estimation (See more detail of across time estimation in Bailey (2007)).

```{r}
# multinomial dynamic
data(sim_match)

# Generating matching bill indicator for across time estimation
bill_match <- make_bill_match(m_mlt_d, sim_match)

dyn_ops_mlt <- make_dyn_options(m_data_dyn,
                                unit_id = "unit",
                                bill_id = "bill",
                                time_id = "time",
                                vote_col = "vote",
                                add_matched_bill = bill_match,
                                clean = TRUE)

fit_mlt_d <- pgIRT(m_mlt_d,
                   mode = "dynamic",
                   constraint = 1,
                   dyn_options = dyn_ops_mlt,
                   verbose = 20)
```

Returning 99% confidence interval:

```{r}
boot_mlt_d <- pgIRT_boot(fit_mlt_d, boot = 100, verbose = 20)

summary(boot_mlt_d, parameter = "theta", ci = .99)
```


## References

+ Bailey, M. A. (2007). "Comparable preference estimates across time and institutions for the court, congress, and presidency". *American Journal of Political Science*, 51(3), 433-448.
+ Goplerud, M. (2019). "A Multinomial Framework for Ideal Point Estimation". *Political Analysis*, 27(1), 69-89.
+ Lewis, J. B., & Poole, K. T. (2004). "Measuring bias and uncertainty in ideal point estimates via the parametric bootstrap". *Political Analysis*, 12(2), 105-127.
+ Martin A.D., Quinn K.M. & Park J.H. (2011). "MCMCpack: Markov Chain Monte Carlo in R." *Journal of Statistical Software*, 42(9), 22.
