---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pgIRT

<!-- badges: start -->
<!-- badges: end -->

`pgIRT` is an **R** package that implements Item Response Theory (IRT) model with Polya-Gamma data augmentation and EM Algorithm. Implementation includes binary, multinomial along with dynamic IRT models. The algorithm here is based on the procedure proposed by Goplerud (2019). This package also utilizes the parametric bootstrap method proposed by Lewis and Poole (2004) to estimate statistical uncertainty. 

## Installation

You can install the released version of pgIRT from [GitHub](https://github.com) with:

``` r
remotes::install_github("vkyo23/pgIRT")
```
## Usage

### Binary IRT

Using `Senate` data from `MCMCpack`.

```{r, warning=FALSE, message=FALSE}
library(pgIRT)
require(dplyr)

data(Senate, package = "MCMCpack")
mat <- as.matrix(Senate[, 6:ncol(Senate)])
colnames(mat) <- 1:ncol(mat)
rownames(mat) <- 1:nrow(mat)

```


### Dynamic binary IRT

### Multinomial IRT

### Dynamic multinomial IRT

## References

- Goplerud, M. (2019). "A Multinomial Framework for Ideal Point Estimation". *Political Analysis*, 27(1), 69-89.
- Lewis, J. B., & Poole, K. T. (2004). "Measuring bias and uncertainty in ideal point estimates via the parametric bootstrap". *Political Analysis*, 12(2), 105-127.
